Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-07-07 introduction.tex
Fixes, refinements in Introduction.

Although the workloads to be executed are comprised of tasks that are
independent of each other, the management of the distribution of workloads
across many heterogeneous resources, the effective utilization of resources
and efficient execution of workloads present non-trivial challenges. Many
software solutions have been developed in response to these challenges. The
CMS experiment, devised a solution based around the
HTCondor~\cite{thain2005distributed} software ecosystem. The
ATLAS~\cite{Aad:2008} experiment utilizes the Production and Distributed
Analysis (PanDA) workload management system~\cite{Maeno2011} (WMS) for
distributed data processing and analysis. The CMS and ATLAS experiments
utilize, arguably the largest production grade distributed computing
solutions and have symbolized the paradigm of {\it high-throughput
computing}, i.e., the effective execution of many independent tasks.

In spite of the impressive scale of the ATLAS distributed computing system,
demand for computing systems will soon significantly outstrip current and
projected supply.   The data volumes that will need analyzing in LHC-Run 3
($~\approx$ 2022) and the high-luminosity era (Run 4) will increase by
factors of 10--100 compared to the current phase (Run 2). There are multiple
levels at which this problem needs to be addressed: the utilization of
emerging parallel architectures (e.g., platforms); algorithmic and advances
in analytical methods (e.g., use of Machine Learning); and the ability to
exploit different platforms (e.g., clouds and supercomputers).

@@

This paper represents the experience of how the ATLAS experiment has ``broken
free'' of the traditional computational approach of high-throughput computing
on distributed resources to embrace new platforms, in particular
high-performance computers. Specifically, we discuss the experience of
integrating PanDA WMS with a US DOE leadership machine (Titan) to reach
sustained production scales of approximately 51M core-hours a year.

We demonstrate how Titan is more efficiently utilized by the mixing of small
and short-lived tasks in backfill with regular payloads. Cycles otherwise
unusable (or very difficult to use) are used for science thus, increasing the
overall utilization on Titan without loss of quality-of-service. The
conventional mix of jobs at OLCF cannot be effectively backfilled because of
size, duration, and scheduling policies. Our approach is extensible to any
HPC with ``capability scheduling'' policies.  We also investigate the use of
a pilot-abstraction based task execution runtime system to flexibly execute
ATLAS and other heterogeneous workloads (molecular dynamics) using regular
queues. As such, our approach provides a general solution and investigation
of the convergence of HPC and HTC execution of workloads.

@@

This paper also provides: (i) a critical evaluation of the many design and
operational considerations that have been taken to support the sustained,
scalable and production usage of Titan for historically high-throughput
workloads, and (ii) early lessons and guidance as the community looks forward
to designing the next generation of online analytical
platforms~\cite{foap-url}, for how current and future experimental and
observational systems can be integrated with production supercomputers in a
general and extensible manner.

>>>>

The LHC workloads are comprised of tasks that are independent of each other,
however, the management of the distribution of workloads across many
heterogeneous resources, the effective utilization of resources and efficient
execution of workloads present non-trivial challenges. Many software solutions
have been developed in response to these challenges. The CMS experiment,
devised a solution based around the HTCondor~\cite{thain2005distributed}
software ecosystem. The ATLAS~\cite{Aad:2008} experiment utilizes the
Production and Distributed Analysis (PanDA) workload management
system~\cite{Maeno2011} (WMS) for distributed data processing and analysis.
The CMS and ATLAS experiments utilize, arguably the largest production grade
distributed computing solutions, and have symbolized the paradigm of {\it
high-throughput computing} (HTC), i.e., the effective execution of many
independent tasks.

In spite of the impressive scale of the ATLAS distributed computing system --
in the number of tasks executed, the number of core hours utilized, and the
number of distributed sites utilized,  demand for computing systems will soon
significantly outstrip current and projected supply.   The data volumes that
will need analyzing in LHC-Run 3 ($~\approx$ 2022) and the high-luminosity era
(Run 4) will increase by factors of 10--100 compared to the current phase (Run
2). There are multiple levels at which this problem needs to be addressed: the
utilization of emerging parallel architectures (e.g., platforms); algorithmic
and advances in analytical methods (e.g., use of Machine Learning); and the
ability to exploit different platforms (e.g., clouds and supercomputers).

@@

This paper represents the experience of how the ATLAS experiment has ``broken
free'' of the traditional computational approach of high-throughput computing
on distributed resources to embrace new platforms, in particular high-
performance computers (HPC). Specifically, we discuss the experience of
integrating PanDA WMS with a US DOE leadership machine (Titan) to reach
sustained production scales of approximately 51M core-hours a year.

In doing so, we demonstrate how Titan is more efficiently utilized by the
mixing of small and short-lived tasks in backfill with regular payloads.
Cycles otherwise unusable (or very difficult to use) are used for science,
thus increasing the overall utilization on Titan without loss of overall
quality-of-service. The conventional mix of jobs at OLCF cannot be effectively
backfilled because of size, duration, and scheduling policies. Our approach is
extensible to any HPC with ``capability scheduling'' policies.  We also
investigate the use of a pilot-abstraction based task execution runtime system
to flexibly execute ATLAS and other heterogeneous workloads (molecular
dynamics) using regular queues. As such, our approach provides a general
solution and investigation of the convergence of HPC and HTC execution of
workloads.

@@

This paper also provides: (i) a critical evaluation of the many design and
operational considerations that have been taken to support the sustained,
scalable and production usage of Titan for historically high-throughput
workloads, and (ii) early lessons and guidance on designing the next
generation of online analytical platforms~\cite{foap-url},  so that
experimental and observational systems can be integrated with production
supercomputers in a general and extensible manner.
